---
output:
  pdf_document: default
  html_document: default
---
### Question 0

*  Draw a concepts diagram that uses all the following Github terminology 
   * Include any missing keywords that will simplify the concepts diagrams 


`Push`, `Repository`, `Clone`, `Pull`, `Pull Request`, `Branch`, `Merging`, `Github Client`, `README file`, `Private or Public`

```{r}
# I cannot figure out how to knit the pdf image to this markdown file. I'll keep trying.
https://www.dropbox.com/s/1t3fu3pqdhrvcfb/Concept_map_2.pdf?dl=1
```

#### Question 1

`matrix(c(1,2,3,4,5,6) , nrow = 3)`

* Running the expression produces the following matrix

|  |  |
|---|---|
| 1 | 4 |
| 2 | 5 |
| 3 | 6 |

* How can you modify the call to `matrix()` to produce the following matrix instead?


|   | col_1 |  col_2 |
| --- |---|---|
| row_1 | 1 | 2 |
| row_2 | 3 | 4 |
| row_3 | 5 | 6 |

* Note that you need to name the columns (col_1 and col_2) and name the rows (row_1, row_2, row_3)

Hint: Use the `?` symbol to invoke the matrix documentation

```{r}
# Write your answer here
test_matrix <- matrix(c(1:6), nrow=3)
rownames(test_matrix)<- c("row_1", "row_2", "row_3")
colnames(test_matrix)<- c("col_1", "col_2")
```

#### Question 2

* Load then sort the airquality data frame on its `Temp` and `Solar.R` columns in reverse order (largest to smallest values)
  * The function to sort a data frame is called order
* Display only the first 15 lines of your table

```{r}
# Write your answer here
airqual_temp<-airquality[order(airquality$Temp, decreasing = TRUE),]
airqual_solar<-airquality[order(airquality$Solar.R, decreasing = TRUE),]
head(airqual_temp, n = 15)
head(airqual_solar, n = 15)
```

#### Question 3

* Sort the airquality data frame on its `Temp` in decreasing order and `Solar.R` in increasing order
* Display only the first 15 lines of your table

```{r}
# Write your answer here
airqual_temp1<-airquality[order(airquality$Temp, decreasing = TRUE),]
airqual_solar1<-airquality[order(airquality$Solar.R, decreasing = FALSE),]
```

#### Question 4

* There are various ways to select a subset of observations from a data frame.

* Consult your [R Reference Card](https://cran.r-project.org/doc/contrib/Baggott-refcard-v2.pdf), see `Data Selection and Manipulation` section.
  * What operations can you use to select all observations where the temperature is 72. Give at least two different answers to this question


```{r}
# Write your answer here
subset(airquality, Temp == 72) #answer 1

temp_72 <- airquality$temp == 72 #answer 2
airquality[temp_72,]
```

#### Question 6

* You may have noticed when working with the `airqulity` data that some values show as `NA`
 * `NA` stands for not available, or missing values.
* A major part of data wrangling consists of cleaning missing values by either:
  * Dropping the lines that have missing values
   * Sometimes we can drop the column with missing values if the column is made of predominantly missing values
  * Imputing the missing values, which uses educated guesses (or more complex algorithms) to fill the missing values
  
* Find and remove all rows that are missing values for the `Solar.R` or `Ozone` variables
* Save the cleaned data to a new data frame called airquality_no_na
  * How many lines have been removed?

```{r}
# Write your answer here

air_quality_no_na<-airquality[is.na(airquality$Solar.R) == FALSE & is.na(airquality$Ozone) == FALSE,]

# 42 lines were removed

```

#### Question 7
* Let's use a different strategy and impute the missing value.
  * replace the missing values for Solar.R using that month's average.
  * Example:
    * The missing value for line 6 should be replaced with the average for month 5.
    * The missing value for line 97 should be replaced with the average for month 8.

```{r}
# Write your answer here
na_solar<-airquality[is.na(airquality$Solar.R),] #subset rows where solar.r is NA
mean_5 <- mean(airquality$Solar.R[airquality$Month %in% unique(na_solar$Month)[1]], na.rm = TRUE) #finds mean of month 5
mean_8 <- mean(airquality$Solar.R[airquality$Month %in% unique(na_solar$Month)[2]], na.rm = TRUE) #finds mean of month 8
copy_airquality <- airquality #because I don't want to mess with the original dataset 
copy_airquality[is.na(airquality$Solar.R[airquality$Month ==5]), 2] <- mean_5 #replaces month 5 na values with mean
copy_airquality[is.na(airquality$Solar.R[airquality$Month ==8]), 2] <- mean_8 #replaces month 8 na values with mean

```

